---
title: "R Notebook"
output: html_notebook
---

Importation des Librairies

```{r}
library(R.matlab)
library(mclust)
library(Rmixmod)
library(aricode)
require(Matrix)
#rankMatrix(M)[1]
```

Sélection de modèle pour un nombre de classes connu
---------------------------------------------------


Importation des Données

```{r}
jaffe.data <- readMat("Dataset/jaffe.mat")

mnist.data <- readMat("Dataset/MNIST5.mat")

mfea.data <- readMat("Dataset/MFEAT1.mat")

usps.data <- readMat("Dataset/USPS.mat")

optdigits.data <- readMat("Dataset/Optdigits.mat")
```


Exploration des Données

```{r}
exploration <- function(data) {
  counts <- table(data$y)
  barplot(counts, main="Number of individus for each classes",
   xlab="classes")
  
  nbsamples=dim(data$X)[1]
  nbfeatures=dim(data$X)[2]
  nbclasses=length(unique(data$y))
  sparsity=1- sum(data$X == 0)/(dim(data$X)[1]*dim(data$X)[2])
  balance=min(table(data$y))/max(table(data$y))
  
  print(paste("# samples", nbsamples))
  print(paste("# features", nbfeatures))
  print(paste("# classes", nbclasses))
  print(paste("sparsity", sparsity))
  print(paste("balance", balance))
  
  return(list(nbsamples=nbsamples, nbfeatures=nbfeatures, nbclasses=nbclasses, sparsity=sparsity, balance=balance))
}

```

```{r}
jaffe.exploration=exploration(jaffe.data)
```
```{r}
mnist.exploraion=exploration(mnist.data)
```
```{r}
mfea.exploration=exploration(mfea.data)
```
```{r}
usps.exploration=exploration(usps.data)
```
```{r}
optdigits.exploration=exploration(optdigits.data)
```
```{r}
exploration.global=data.frame(rbind(jaffe.exploration, optdigits.exploration, mfea.exploration, usps.exploration, mnist.exploraion), row.names=c("jaffe", "optdigits", "mfea", "mnist", "usps"))
```


1. ALL POSSIBLE MIXTURE MODEL

```{r}
mixt_mod <- function(data) {
  models_partition <- c()
  loglike <- c()
  bic <- c()
  models<-c()
  for(model in mclust.options("emModelNames")){
    #cat("=======\n# ", model, "\n=======\n")
    res_temp <- try(Mclust(data, G = 10, modelNames = model,verbose = F ))
    if(typeof(res_temp) == "list"){
      to_add <- res_temp$classification
      models_partition <- cbind(models_partition, to_add)
      loglike <- c(loglike, res_temp$loglik)
      bic <- c(bic, res_temp$bic)
      models <- c(models, model)
      #print("!!! SUCESS !!!")
    }else{
      print(paste("!!!",model, "FAILED !!!"))
    }
  }
  return(list(models_partition=models_partition, loglike=loglike, bic=bic, models=models))
}
    
```
```{r}

```

```{r}
jaffe.svd=svd(jaffe.data$X)
ind=max(which(jaffe.svd$d>1))
D <- diag(jaffe.svd$d[1:ind])
jaffe.mixt_mod=mixt_mod(jaffe.data$X %*% jaffe.svd$v[,1:ind])
```

```{r}
print("Liste des modeles :")
print(jaffe.mixt_mod$models)
```


```{r}
optdigits.svd=svd(optdigits.data$X)
ind=max(which(optdigits.svd$d>1))
D <- diag(optdigits.svd$d[1:ind])
optdigits.mixt_mod=mixt_mod(optdigits.data$X %*% optdigits.svd$v[,1:ind])
```
```{r}
print("Liste des modeles :")
print(optdigits.mixt_mod$models)
```


```{r}
mfea.svd=svd(mfea.data$X)
ind=max(which(mfea.svd$d>1))
D <- diag(mfea.svd$d[1:ind])
mfea.mixt_mod=mixt_mod(mfea.data$X %*% mfea.svd$v[,1:ind])
```

```{r}
print("Liste des modeles :")
print(mfea.mixt_mod$models)
```



2. BEST MIXTURE MODEL

```{r}
jaffe.best_mixt_mod <- Mclust(jaffe.data$X, G = 10,verbose = F )
jaffe.best_mixt_mod.NMI=NMI(as.factor(jaffe.best_mixt_mod$classification), as.factor(jaffe.data$y))
print(paste("NMI FOR BEST MIXTURE MODEL :", jaffe.best_mixt_mod.NMI))
```
```{r}
optdigits.best_mixt_mod <- Mclust(optdigits.data$X, G = 10,verbose = F )rbinrbin
optdigits.best_mixt_mod.NMI=NMI(as.factor(optdigits.best_mixt_mod$classification), as.factor(optdigits.data$y))
print(paste("NMI FOR BEST MIXTURE MODEL :", optdigits.best_mixt_mod.NMI))
```

```{r}
mfea.best_mixt_mod <- Mclust(mfea.data$X, G = 10,verbose = F )
mfea.best_mixt_mod.NMI=NMI(as.factor(mfea.best_mixt_mod$classification), as.factor(mfea.data$y))
print(paste("NMI FOR BEST MIXTURE MODEL :", mfea.best_mixt_mod.NMI))
```

```{r}
resultats.best_mixt_mod=data.frame("NMI"=rbind(jaffe.best_mixt_mod.NMI, optdigits.best_mixt_mod.NMI, mfea.best_mixt_mod.NMI))
```

3. CONSENSUS OVER ALL MIXTURE MODEL

```{r}
cluster_consensus <- function(partitions, label) {
  consensus_tmp=mixmodCluster(as.data.frame(type.convert(partitions)), nbCluster = 10, models=mixmodMultinomialModel(), dataType="qualitative")
  NMI_tmp=NMI(as.factor(consensus_tmp@bestResult@partition), as.factor(label))
  print(paste("NMI FOR CONSENSUS OVER ALL MIXTURE MODEL :", NMI_tmp))
  return(list(NMI=NMI_tmp, consensus=consensus_tmp))
}
```

```{r}
jaffe.consensus=cluster_consensus(jaffe.mixt_mod$models_partition, jaffe.data$y)
```
```{r}
optdigits.consensus=cluster_consensus(optdigits.mixt_mod$models_partition, optdigits.data$y)
```


```{r}
mfea.consensus=cluster_consensus(mfea.mixt_mod$models_partition, mfea.data$y)
```

```{r}
resultats.consensus=data.frame("NMI"=rbind(jaffe.consensus$NMI,optdigits.consensus$NMI, mfea.consensus$NMI))
```


4. CONSENSUS OVER BEST MIXTURE MODEL

rather than doing a consensus over all the mixture model, we can do a consensus over 2 the best mixture model in terms of log likelihood score and BIC score

```{r}
jaffe.custom_consensus_log_lik=cluster_consensus(jaffe.mixt_mod$models_partition[,head(order(jaffe.mixt_mod$loglike),3)], jaffe.data$y)
jaffe.custom_consensus_bic=cluster_consensus(jaffe.mixt_mod$models_partition[,head(order(jaffe.mixt_mod$bic),3)], jaffe.data$y)
```
```{r}
optdigits.custom_consensus_log_lik=cluster_consensus(optdigits.mixt_mod$models_partition[,head(order(optdigits.mixt_mod$loglike),3)], optdigits.data$y)
optdigits.custom_consensus_bic=cluster_consensus(optdigits.mixt_mod$models_partition[,head(order(optdigits.mixt_mod$bic),3)], optdigits.data$y)
```


```{r}
mfea.custom_consensus_log_lik=cluster_consensus(mfea.mixt_mod$models_partition[,head(order(mfea.mixt_mod$loglike),2)], mfea.data$y)
mfea.custom_consensus_bic=cluster_consensus(mfea.mixt_mod$models_partition[,head(order(mfea.mixt_mod$bic),2)], mfea.data$y)
```

```{r}
resultats.custom_consensus_log_lik=data.frame("NMI"=rbind(jaffe.custom_consensus_log_lik$NMI,optdigits.custom_consensus_log_lik$NMI, mfea.custom_consensus_log_lik$NMI))
resultats.custom_consensus_bic=data.frame("NMI"=rbind(jaffe.custom_consensus_bic$NMI,optdigits.custom_consensus_bic$NMI, mfea.custom_consensus_bic$NMI))
```


```{r}
resultats.global=data.frame(cbind(resultats.best_mixt_mod, resultats.consensus, resultats.custom_consensus_bic, resultats.custom_consensus_log_lik), row.names=c("jaffe", "optdigits", "mfea"))
colnames(resultats.global)=c("best", "consensus", "consensusbic", "consensusloglik")
resultats.global
```

```{r}
#write.csv(exploration.global, file="exploration_global.csv")
write.csv(resultats.best_mixt_mod, file = 'resultats_best_mixt_mod.csv')
write.csv(resultats.consensus, file = "resultats_consensus.csv")
write.csv(resultats.custom_consensus_bic, file = "resultats.custom_consensus_bic.csv")
write.csv(resultats.custom_consensus_log_lik, file="resultats.custom_consensus_log_lik.csv")
write.csv(resultats.global, file="resultats_global.csv")
```


